Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-27 background.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.

% TODO: Change these pictures to match look-n-feel
\begin{figure}[ht]
\centering
\subfigure[Centralised (SETI@Home or Napster).] {
  \includegraphics[scale=0.3]{img/centralised.png}
  \label{figure:p2p-archs:centralised}
}\qquad\qquad
\subfigure[Fully unstructured (Gnutella v.0.4).] {
  \includegraphics[scale=0.3]{img/unstructured-full.png}
  \label{figure:p2p-archs:unstructured-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
  \includegraphics[scale=0.3]{img/unstructured-hierarchical.png}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad\qquad
\subfigure[Structured (Chord and Kademlia).] {
  \includegraphics[scale=0.3]{img/structured.png}
  \label{figure:p2p-archs:structured}
}
\caption{The various peer-to-peer architectures.}
\label{figure:p2p-archs}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[scale=0.8]{img/phys.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}

@@

\begin{figure}[ht]
\centering
\subfigure[Nodes B and C have direct overlay connection.] {
  \includegraphics[scale=0.8]{img/overlay-1.pdf}
  \label{figure:overlay-1}
}\qquad\qquad
\subfigure[Nodes A and D have direct overlay connection.] {
  \includegraphics[scale=0.8]{img/overlay-2.pdf}
  \label{figure:overlay-2}
}
\caption{Two different overlay connection configuration.}
\label{figure:overlay-confs}
\end{figure}

>>>>

% TODO: Change these pictures to match look-n-feel
\begin{figure}[ht]
\centering
\subfigure[Centralised (SETI@Home or Napster).] {
  \includegraphics[scale=0.25]{img/pdf/centralised.pdf}
  \label{figure:p2p-archs:centralised}
}\qquad\qquad
\subfigure[Fully unstructured (Gnutella v.0.4).] {
  \includegraphics[scale=0.25]{img/pdf/unstructured-full.pdf}
  \label{figure:p2p-archs:unstructred-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
  \includegraphics[scale=0.25]{img/pdf/unstructured-hierarchical.pdf}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad\qquad
\subfigure[Structured (Chord and Kademlia).] {
  \includegraphics[scale=0.25]{img/pdf/structured.pdf}
  \label{figure:p2p-archs:structured}
}
\caption{The various peer-to-peer architectures.}
\label{figure:p2p-archs}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[scale=0.7]{img/pdf/example-physical.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}

@@

\begin{figure}[ht]
\centering
\subfigure[Nodes B and C have direct overlay connection.] {
  \includegraphics[scale=0.5]{img/pdf/example-overlay-inefficient.pdf}
  \label{figure:overlay-1}
}\qquad\qquad
\subfigure[Nodes A and D have direct overlay connection.] {
  \includegraphics[scale=0.5]{img/pdf/example-overlay-efficient.pdf}
  \label{figure:overlay-2}
}
\caption{Two different overlay connection configuration.}
\label{figure:overlay-confs}
\end{figure}

### 2014-01-27 structured.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.

\begin{figure*}
\centering
  \includegraphics[scale=0.8]{img/algorithms/global_softstate}
\caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position
of $p$ on the map.}
\label{fig:global_softstate}
\end{figure*}

\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. Figure \ref{fig:global_softstate} presents the landmark
space and the mapping of the position of the nodes. This global view of the
state is made available to all nodes in order to help them find the best way to
route their messages. \textit{Global Soft-State} operates in two main stages,
generation and use of the proximity information. For generating the proximity
information a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step in order to select a number of potential nearest neighbour
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm chooses a different path from the
classic gossiping approaches for constructing and maintaining the overlay. It is
based on landmark clustering for strategic placement of proximity information on
the overlay enabling any node to access such information using a landmark number
that reflects its physical position in the network. For various logical
regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
context or a set of nodes sharing a particular prefix in overlays such as
Pastry.} maps of physical information are built and published where each node
may appear in a maximum of $log\left( N \right)$ such maps. To dynamically adapt
to changing network conditions, a node subscribes to relevant \emph{soft states}
that utilize a notification system in order to initialize any necessary
neighbour re-selection.

@@

\begin{figure}[ht]
\centering
\subfigure[Example physical topology of nodes.] {
  \includegraphics[scale=0.6]{img/algorithms/SAT_match}
  \label{figure:sat_match:before}
}\qquad\qquad
\subfigure[(left) A jumps to B in order to match the logical topology to the
physical. (right) State of the logical topology after the jump.] {
  \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
  \label{figure:sat_match:after}
}
\caption{The SAT-Match algorithm.}
\label{figure:sat_match}
\end{figure}

\emph{SAT-Match} \cite{RGJZ2004}. Even though no theoretical proof is
given, valid results are obtained and demonstrated through experimentations.
Similar to the other approaches in the proximity neighbour selection,
\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
information. However, as an additional feature, SAT-Match peers can do
selective jumps to adjust their positioning in the DHT, if it reduces the
\emph{stretch} of its one-hop neighbourhood. The paper defines stretch
$S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average
logical link latency while the $\bar{L_p}$ is the average physical
link latency, and uses it a way quantify the \emph{topology
match degree} of the constructed overlay. For the probing phase SAT-Match uses a
small TTL value for the probing messages in order to reduce
redundancy \cite{jiang_lightflood_2008}. This process begins as soon as the node
joins the network using a DHT mechanism. Each probing message contains
information about the source and a small TTL value. The recipient of such a
message returns information about itself to the source and forwards the probing
message to its neighbours if the TTL is non-zero. The discovered nodes are
referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$
distance to the source node. Blindly selecting the peer with the smallest RTT
as neighbour is, generally, not the optimal decision to make in order to achieve
global stretch reduction. In a structured scheme, when a node jumps to connect
to a physically close node, it may need to connect to other distant nodes to
maintain the structure's integrity, thus creating an overall increase in the
overlay's stretch. The two nodes with the smallest RTT are then used in order to
select one zone to jump in this phase. The algorithm is as follows: The source
node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
the $TTL-1$ neighbourhood of the first of the previously selected peers. These
calculations are made as if the jump has been made. If the stretch reduction is
over a predefined threshold the jump is performed, otherwise the second selected
candidate is picked and the same computations are performed. If again, the
threshold is not met, then no jump is ultimately done. In case of a jump, this
is performed as a combination of \emph{leave} and \emph{join} operations, in the
CAN context. Figure~\ref{figure:sat_match} demonstrates the selective jump on an
example topology.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{ \bf Delay Aware P2P System}
A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
main goal is to reduce the time $L$ of a look-up request by dividing the routing
tables of peers into several sectors in increasing delay. The source node
emitting the query message defines a delay boundary or the pruning factor $L_t$
in the paper's context. Request messages will be forwarded only to nodes whose
delay less than or equal to $L_t$. With the clustered routing tables and the
loose organisation the overlay network of \emph{DAPS} is between structured and
unstructured.

% TODO: MAYBE REVIEW THE EVALUATION
\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
low &
%
low &
%
medium
\end{tabular}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Geographic layout} \label{section:geographic_layout}
Inside geographic layout category fall those kind of methods which practically
position physically close by nodes together in the application space as well.
This is done by adjusting and maintaining the routing tables of all involved
peers accordingly by exploiting proximity information that describe the overall
geographic positioning of the peer nodes. Landmark servers and RTT measurements
are two popular methods, which can also be used in conjunction, to discover
physically close by peers over the network. However, as discussed in Section
\ref{sec:landmark_binning}, these methods do not always give reliable estimates
for the node positions over the internet. The landmark servers are not
self-organizing and have maintenance overheads. To serve a P2P network with
millions of peers, multiple landmark servers distributed over the whole world is
required, which is hard to manage. The RTT measurements also can measure the
delay between peers, but it is a greedy method, which can result non-optimal
overlay topologies especially if close by nodes have low bandwidth connections
among themselves.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proximity Routing}
Proximity routing does not require routing tables to be built using any
knowledge about network proximity. On the other hand it exploits such knowledge
in order to choose the best next hop during routing a message. This approach
tries to balance between choosing the node that will further progress the
routing towards the destination and choosing the closest entry in the routing
table, in terms of network proximity. As this is very difficult to be achieved,
the system might end up using a greedy approach that selects low latency paths
over the overlay, which maps to suboptimal longer paths on the physical
topology. Thus, in general, proximity neighbour selection is considered as a
superior method to proximity routing, however, joint uses of these two protocols
are also possible. Moreover, they are considered relatively less effective than
geographical layout when applied to CAN(-like) implementations. Moreover, the
technique has been incorporated into a version of Chord causing an increase on
the overhead of node joins and the size as well as maintenance cost of finger
tables.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proximity neighbour selection}
\emph{Proximity Neighbour Selection} constructs the routing tables using
proximity knowledge. The proximity information used in this method is different
than the landmark based systems described in
Section~\ref{section:geographic_layout}, as time-to-live values between nodes,
or directly node ID prefixes are used to detect proximity. Tapestry, Pastry, and
CAN successfully implemented the proximity into their algorithms by using this
approach. The routing protocol in Pastry is based on longest node ID prefix
matching, while CAN uses RTT values to detect close by nodes.
\cite{CDCR2002a} reports that proximity neighbour selection as an
effective proximity based method.

>>>>

% \begin{figure*}
% \centering
%   \includegraphics[scale=0.8]{img/algorithms/global_softstate}
% \caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position
% of $p$ on the map.}
% \label{fig:global_softstate}
% \end{figure*}

\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state is made available to all nodes in order to help them find the best way to
route their messages. \textit{Global Soft-State} operates in two main stages,
generation and use of the proximity information. For generating the proximity
information a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step in order to select a number of potential nearest neighbour
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm chooses a different path from the
classic gossiping approaches for constructing and maintaining the overlay. It is
based on landmark clustering for strategic placement of proximity information on
the overlay enabling any node to access such information using a landmark number
that reflects its physical position in the network. For various logical
regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
context or a set of nodes sharing a particular prefix in overlays such as
Pastry.} maps of physical information are built and published where each node
may appear in a maximum of $log\left( N \right)$ such maps. To dynamically adapt
to changing network conditions, a node subscribes to relevant \emph{soft states}
that utilize a notification system in order to initialize any necessary
neighbour re-selection.

@@

% \begin{figure}[ht]
% \centering
% \subfigure[Example physical topology of nodes.] {
%   \includegraphics[scale=0.6]{img/algorithms/SAT_match}
%   \label{figure:sat_match:before}
% }\qquad\qquad
% \subfigure[(left) A jumps to B in order to match the logical topology to the
% physical. (right) State of the logical topology after the jump.] {
%   \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
%   \label{figure:sat_match:after}
% }
% \caption{The SAT-Match algorithm.}
% \label{figure:sat_match}
% \end{figure}

\emph{SAT-Match} \cite{RGJZ2004}. Even though no theoretical proof is
given, valid results are obtained and demonstrated through experimentations.
Similar to the other approaches in the proximity neighbour selection,
\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
information. However, as an additional feature, SAT-Match peers can do
selective jumps to adjust their positioning in the DHT, if it reduces the
\emph{stretch} of its one-hop neighbourhood. The paper defines stretch
$S = \frac{\bar{L_l}}{\bar{L_p}}$, where $\bar{L_l}$ is the average
logical link latency while the $\bar{L_p}$ is the average physical
link latency, and uses it a way quantify the \emph{topology
match degree} of the constructed overlay. For the probing phase SAT-Match uses a
small TTL value for the probing messages in order to reduce
redundancy \cite{jiang_lightflood_2008}. This process begins as soon as the node
joins the network using a DHT mechanism. Each probing message contains
information about the source and a small TTL value. The recipient of such a
message returns information about itself to the source and forwards the probing
message to its neighbours if the TTL is non-zero. The discovered nodes are
referred to as $TTL-k$ neighbourhood of the source node based on the $TTL$
distance to the source node. Blindly selecting the peer with the smallest RTT
as neighbour is, generally, not the optimal decision to make in order to achieve
global stretch reduction. In a structured scheme, when a node jumps to connect
to a physically close node, it may need to connect to other distant nodes to
maintain the structure's integrity, thus creating an overall increase in the
overlay's stretch. The two nodes with the smallest RTT are then used in order to
select one zone to jump in this phase. The algorithm is as follows: The source
node $S$ calculates the stretch change of its $TTL-1$ neighbourhood and that of
the $TTL-1$ neighbourhood of the first of the previously selected peers. These
calculations are made as if the jump has been made. If the stretch reduction is
over a predefined threshold the jump is performed, otherwise the second selected
candidate is picked and the same computations are performed. If again, the
threshold is not met, then no jump is ultimately done. In case of a jump, this
is performed as a combination of \emph{leave} and \emph{join} operations, in the
CAN context.
% Figure~\ref{figure:sat_match} demonstrates the selective jump on an
% example topology.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Geographic layout} \label{section:geographic_layout}

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/geographic-layout.pdf}
\caption{Successive nodes in the overlay circle can be chosen to be from the
same or near autonomous systems.}
\label{figure:geographic-layout}
\end{figure}

Inside geographic layout category fall those kind of methods which practically
position physically close by nodes together in the application space as well
(check Figure~\ref{figure:geographic-layout}). This is done by adjusting and
maintaining the routing tables of all involved peers accordingly by exploiting
proximity information that describe the overall geographic positioning of the
peer nodes. Landmark servers and RTT measurements are two popular methods, which
can also be used in conjunction, to discover physically close by peers over the
network. However, as discussed in Section~\ref{sec:landmark_binning}, these
methods do not always give reliable estimates for the node positions over the
internet. The landmark servers are not self-organizing and have maintenance
overheads. To serve a P2P network with millions of peers, multiple landmark
servers distributed over the whole world is required, which is hard to manage.
The RTT measurements also can measure the delay between peers, but it is a
greedy method, which can result in non-optimal overlay topologies especially if
close by nodes have low bandwidth connections among themselves.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proximity Routing}

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
\caption{Proximity routing takes proximity info into account during the
routing process.}
\label{figure:proximity-routing}
\end{figure}

Proximity routing does not require routing tables to be built using any
knowledge about network proximity. On the other hand it exploits such knowledge
in order to choose the best next hop during routing a message as can be seen in
Figure~\ref{figure:proximity-routing}. This approach tries to balance between
choosing the node that will further progress the routing towards the destination
and choosing the closest entry in the routing table, in terms of network
proximity. As this is very difficult to be achieved, the system might end up
using a greedy approach that selects low latency paths over the overlay, which
maps to suboptimal longer paths on the physical topology. Thus, in general,
proximity neighbour selection is considered as a superior method to proximity
routing, however, joint uses of these two protocols are also possible. Moreover,
they are considered relatively less effective than geographical layout when
applied to CAN(-like) implementations. Moreover, the technique has been
incorporated into a version of Chord causing an increase on the overhead of node
joins and the size as well as maintenance cost of finger tables.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proximity neighbour selection}

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/proximity-neighbour-selection.pdf}
\caption{Proximity neighbour selection dictates that proximity is taken into
acount during routing table maintenance.}
\label{figure:proximity-neighbour-selection}
\end{figure}

\emph{Proximity Neighbour Selection} constructs the routing tables using
proximity knowledge (see Figure~\ref{figure:proximity-neighbour-selection}). The
proximity information used in this method is different from the the
landmark-based systems described in Section~\ref{section:geographic_layout}.
Here time-to-live values between nodes, or directly node ID prefixes are used to
detect proximity. Tapestry, Pastry, and CAN successfully implemented the
proximity into their algorithms by using this approach. The routing protocol in
Pastry is based on longest node ID prefix matching, while CAN uses RTT values to
detect close by nodes. \cite{CDCR2002a} reports that proximity neighbour
selection as an effective proximity based method.
